---
title: "Bank Analysis And Modeling"
author: "Taylor Bonar & Michael Burgess & Rashmi Patel"
date: "7/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(naniar)
library(GGally)
library(funModeling)
library(Hmisc)
setwd(".")
basic_eda <- function(data) # Sample Function Source: https://blog.datascienceheroes.com/exploratory-data-analysis-in-r-intro/
{
  glimpse(data)
  print(status(data))
  freq(data) 
  print(profiling_num(data))
  plot_num(data)
  describe(data)
}
```
# Introduction
The goal of this paper is to investigate banking data via an exploratory data analysis. Once we have initially examined the data, we will then move forward with attempting a classification model via logistic regression for predicting whether or not a client will subscribe to a banking institution given a direct marketing campaign taking place.

# Exploratory Data Analysis

The data we'll be exploring is from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Bank+Marketing) relating to the direct marketing campaigns of a Portuguese banking institution. The initial data set consists of 45,211 observations with 20 inputs, dating between May 2008 to November 2010.

As described by the UCI Machine Learning Repository, each of the variables/columns are described as:

>**Input variables:**
>**bank client data:**
>
>1 - age (numeric)
>
>2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
>
>3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
>
>4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
>
>5 - default: has credit in default? (categorical: 'no','yes','unknown')
>
>6 - housing: has housing loan? (categorical: 'no','yes','unknown')
>
>7 - loan: has personal loan? (categorical: 'no','yes','unknown')
>
>**related with the last contact of the current campaign:**
>
>8 - contact: contact communication type (categorical: 'cellular','telephone')
>
>9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
>
>10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
>
>11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
>
>**other attributes:**
>
>12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
>
>13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
>
>14 - previous: number of contacts performed before this campaign and for this client (numeric)
>
>15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
>
># social and economic context attributes
>
>16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
>
>17 - cons.price.idx: consumer price index - monthly indicator (numeric)
>
>18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
>
>19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
>
>20 - nr.employed: number of employees - quarterly indicator (numeric)
>
>**Output variable (desired target):**
>
>21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

Another caution we will need to access is that within this dataset, there is potential for more than one contact to the same client as it was required to access the product, a bank term deposit would or would not be subscribed.

```{r EDA}
# Retrieve datasets zip
if(!file.exists("./data/bank.zip")) {
  download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/00222/bank-additional.zip", "./bank-additional.zip", mode="wb")
}
unzip("./bank-additional.zip", files = c("bank-additional/bank-additional.csv","bank-additional/bank-additional-full.csv","bank-additional/bank-additional-names.txt"))

# Read data into a data frame object
full_bank <- read.csv("./bank-additional/bank-additional-full.csv", header = T, sep = ";")

# Use function to create initial data insights for bank data
basic_eda(full_bank)

# Transform all chr objects in data frame to a factor class as a secondary data frame object
full_bank_2 <- as.data.frame(unclass(full_bank), stringsAsFactors = T)
```


