---
title: "Bank 6372 Project 2"
author: "Rashmi Patel"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(ellipsis)
library(tidyverse)
library(corrplot)
library(mlbench)
library(caret)
library(skimr)
library(mice)
library(purrr)
library(ggplot2)
library(ggthemes)
library(cowplot)
library(class)
library(e1071)
library(grid)
library(tidyr)
library(stringr)
library(naniar)
library(car)
library(MASS)
library(caret)
library(ROCR)
library(glmnet)
library(bestglm)
library(car)
library(ResourceSelection)
library(randomForest)
library('SmartEDA')
library(kableExtra)
library(DataExplorer)
```
# Load the data from Github 
```{r}
bank=read_delim("https://raw.githubusercontent.com/RashmiAPatel19/SMU-6372-Applied-Stats/main/Project%202/data/bank-additional-full.csv", ";", escape_double = FALSE, trim_ws = TRUE)
bank
```
# Addressing Summary of the Bank data
```{r}
head(bank)
tail(bank)
str(bank)
dim(bank)
names(bank)
summary(bank)
```
# Addressing missing values

* There is no missing values

```{r}
# Checking for missing values
sum(is.na(bank))

```

# Addressing Character Variable and Numeric Variable 
```{r}
# Checking the dimensions of the dataset
dim(bank)
# Checking the column names of the dataset
colnames(bank)
# Looking at the summary of the dataset
summary(bank)
# Checking for data types of the columns of the dataset
str(bank)
# Checking for number of columns with numeric type
numeric_var_who=sum(sapply(bank[,1:21],is.numeric))
numeric_var_who
# Checking for number of columns with character type
char_var_who=sum(sapply(bank[,1:21],is.character))
char_var_who
# Checking for column names with numeric type
numeric_varname_who=which(sapply(bank[,1:21],is.numeric))
numeric_varname_who
# Checking for column names with character type
char_varname_who=which(sapply(bank[,1:21],is.character))
char_varname_who
```

# Looking at the normality of Numeric Variables
```{r}

plot_histogram(bank)

```

# Looking at the graph for Response variable
```{r}

ggplot(bank,aes(x=y,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))

```
# 
```{r}
# Age vs Response
ggplot(bank,aes(x=age,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)
# Job vs Response
ggplot(bank,aes(x=job,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle=90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Marital Status vs Response
ggplot(bank,aes(x=marital,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Education vs Response
ggplot(bank,aes(x=education,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Housing vs Response
ggplot(bank,aes(x=housing,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Default vs Response
ggplot(bank,aes(x=default,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Loan vs Response
ggplot(bank,aes(x=loan,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Contact vs Response
ggplot(bank,aes(x=contact,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Day of the week vs Response
ggplot(bank,aes(x=day_of_week,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Duration vs Response
ggplot(bank,aes(x=duration,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Campaign vs Response
ggplot(bank,aes(x=campaign,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Pdays vs Response
ggplot(bank,aes(x=pdays,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Previous vs Response
ggplot(bank,aes(x=previous,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Poutcome vs Response
ggplot(bank,aes(x=poutcome,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Emp.var.Rate vs Response
ggplot(bank,aes(x=emp.var.rate,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Cons.price.idx vs Response
ggplot(bank,aes(x=cons.price.idx,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Cons.conf.idx vs Response
ggplot(bank,aes(x=cons.cond.idx,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Euribor3m vs Response
ggplot(bank,aes(x=euribor3m,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# nr.employed vs Response
ggplot(bank,aes(x=nr.employed,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))
# Month vs Response
ggplot(bank,aes(x=month,fill=y))+geom_bar(position="stack",stat="count")+
  theme(axis.text.x = element_text(angle = 90))+facet_wrap(~y)+
  geom_text(aes(label=..count..),stat="count",position=position_stack(0.5))


```
This table shows the correlation between the numerical variables

 - nr.employed and emp.var.rate are 91% correlated. 
 - nr.employed and euribor3m are 95% correlated.
 - emp.var.rate and euribor3m are 97% correlated.
 - cons.price.idx and emp.var.rate are 78% correlated.
 - cons.price.idx and euribor3m are 69% correlated.
 - cons.price.idx and nr.employed are 52% correlated.
 
Later we will examine pairwise multicolinearity within the continuous explanatory variables and VIF to see if which explanatory variables may be redundant.

```{r}
corrdfTraintable <- bank %>% keep(is.numeric) %>% na.omit %>% cor %>% view

bank %>% keep(is.numeric) %>% na.omit %>% cor %>% corrplot("upper", addCoef.col = "black", number.digits = 2, number.cex = 0.5, method="shade", order = "hclust", tl.srt=45, tl.cex = 0.8)

view(corrdfTraintable)

```
We don't observe any multicolinearity within the numeric responses, and don't see any need to remove any based on pairwise comparison. 

```{r}

my.cor <- cor(bank %>% keep(is.numeric) %>% na.omit)

library(gplots)
library(ggplot2)
heatmap.2(my.cor,col=redgreen(75), 
          density.info="none", trace="none", dendrogram=c("row"), 
          symm=F,symkey=T,symbreaks=T, scale="none")


#Another option here would be to do PCA among the continous predictors to see
#if they seperate out.  Or a heatmap.
pc.result<-prcomp(bank %>% keep(is.numeric) %>% na.omit,scale.=TRUE)
pc.scores<-pc.result$x
pc.scores<-data.frame(pc.scores)
pc.scores$y<-bank$y


#Use ggplot2 to plot the first few pc's
ggplot(data = pc.scores, aes(x = PC1, y = PC2)) +
  geom_point(aes(col=y), size=1)+
  ggtitle("PCA of y")

ggplot(data = pc.scores, aes(x = PC2, y = PC3)) +
  geom_point(aes(col=y), size=1)+
  ggtitle("PCA of y")

ggplot(data = pc.scores, aes(x = PC3, y = PC4)) +
  geom_point(aes(col=y), size=1)+
  ggtitle("PCA of y")

# we don't really see much separation with PCA

```
Missing values are disguised as unknown values in the data set and we devised a code to show them all

 default           8597
 education         1731
 housing            990
 loan               990
 job                330
 marital             80
 
 For Marital, housing and job it is safe to remove the unknown values as they're so little they won't have an effect on the rest of the distribution
 
 * Looking at default is has no information as it is highly skewed towards "no" as only 3 counts comes up as "yes" so we are removing "default" as well.
 
 
 
 
```{r}
bank %>% 
  summarise_all(list(~sum(. == "unknown"))) %>% 
  gather(key = "variable", value = "nr_unknown") %>% 
  arrange(-nr_unknown)

summary(bank$job)
summary(bank$default)
summary(bank$education)
summary(bank$loan)
summary(bank$marital)
summary(bank$housing)

bank <- subset(bank, job!="unknown")
bank <- subset(bank, marital!="unknown")
bank <- subset(bank, housing!="unknown")
bank <- subset(bank, loan!="unknown")
bank <- subset(bank, default!="unknown")
bank <- subset(bank, education!="unknown")


```
```{r}

library("dplyr")
ShowPieChart <- function(columnBy, columnToShow, titleName)
{
  df <- dplyr::group_by(bank, .dots = c(columnBy, columnToShow)) %>%
    dplyr::summarize(counts = n()) %>%
    dplyr::mutate(perc = (counts / sum(counts)) * 100) %>%
    dplyr::arrange_(.dots=c(columnBy, columnToShow))
 
  
  # preparing the plot
  ggplot2::ggplot(df, aes('', counts)) +
    geom_col(
      position = 'fill',
      color = 'black',
      width = 1,
      aes(fill = y)
    ) +
    ggtitle(titleName) +
    facet_wrap(paste("~",columnBy), labeller = "label_both") +
    geom_label(
      aes(label = paste0(round(perc), "%"), group = "y"),
      position = position_fill(vjust = 0.5),
      color = 'black',
      size = 5,
      show.legend = FALSE
    ) + scale_fill_discrete(name = "Outcome:") +
    coord_polar(theta = "y")
}
ShowPieChart("job", "y", "Outcome by Job")
ShowPieChart("marital", "y", "Outcome by Marital Status")
ShowPieChart("education", "y", "Outcome by Education")
ShowPieChart("housing", "y", "Outcome by Housing")
ShowPieChart("default", "y", "Outcome by Credit In Default")
ShowPieChart("loan", "y", "Outcome by loan status")
ShowPieChart("contact", "y", "Outcome by Contact")
ShowPieChart("poutcome", "y", "Outcome by poutcome")
 
```
